\begin{center}
{\Huge \bf{Abstract}\vskip 0.2in}
\end{center}

\noindent A messaging app is a big undertaking. Making it needs the collaboration of many technologies and computing disciplines. This report represents the background study, technology used development, implementation and testing of a messaging app called Paigaam.Paigaam is a multiplatform app which includes an android app and a webapp. It features chatting through XMPP(Extensible Messaging and Presence Protocol) and the chat is end-to-end encrypted. Users can make their own groups and send images to each other.\\

\noindent The unique thing about this app is that it also features a chatbot called Paigam-Assistant that can respond to multiple queries from students and teachers. The students can ask frequently asked questions such as attendance details in subjects, fees paid status, enrolled courses, teachers details, student details, average percentage in semesters to the bot and get responses.\\

\noindent The App is the one stop solution for all communication needs for the students and faculty members inside and out of the college campus. It is a full-fledged messaging App offering a complete, rich and innovative feature set. Users don't have to suffer through any enraging registration. Log in using your college ID and password and explore the college centric social network.
\\Here are some of the App's most lavishing features:
\begin{itemize}
\item Chat using multiple accounts from a single copy of the App.
\item Encrypt your messages over the medium.
\item Automatic conversation backup on the server, sparing your precious phone storage.
\item Groups for conference chats.
\item Share files and documents of any type such as pdf, docx, images stc.
\item Save your contacts on your phone and / or with your account and import them to any phone via login.
\item Video chat through the webapp.

\end{itemize}